Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations

### 2024-08-13 Delaneyetal2024.tex
pulling from OL

@@

The second experiment aims to characterize the variability in sediment discharge capacity in subglacial relative to subaerial channels across a range of channel slopes and shapes, and friction factors.
Additionally, we examine the effects of differing water discharge smoothing periods from $15$\,\unit{min} up to $15$\,\unit{days}.
The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
For brevity, these values together are referred to as ``model outputs'' in the text.
To evaluate variability, we subtract the model outputs from their daily averages. 
This creates a time series that is detrended from he seasonal variations in the model outputs and has an approximately normal distribution.
We then present the variability in the model output by taking the standard deviation of this detrended time series.

@@

@@

Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different channel types and for different $\alpha$.
Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow channel type.
The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$ for the MPM (Equation~\ref{eq:Qs_eq}).
Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used previously scale only as $Q^2$ for pipe-flow, whereas the sediment transport capacity scales at least with $Q^3$.
The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
However, for the value of $\alpha$ around $\frac{1}{3}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.

>>>>

\ian{include figure}

@@

The second experiment aims to characterize the variability in sediment discharge capacity in subglacial relative to subaerial channels across a range of channel slopes and shapes, and friction factors.
Additionally, we examine the effects of differing water discharge smoothing periods from $15$\,\unit{min} up to $15$\,\unit{days}.
The results below present water velocity ($v_g$, $v_f$), shear stress ($\tau_g$, $\tau_f$), and width-integrated shear stress ($w_g\tau_g$, $w_f\tau_f$) from the subglacial and subaerial models.
For brevity, these values together are referred to as ``model outputs'' in the text.
To evaluate variability, we subtract the model outputs from their daily averages. 
This creates a time series that is detrended from the seasonal variations in the model outputs and has an approximately normal distribution.
We then present the variability in the model output by taking the standard deviation of this detrended time series.

@@

\ian{is this figure better? ... text reorganized. What about a symbol for cross-referencing}

@@

Table~\ref{tab:Qs} shows the scaling of the proxy $w\tau$ as well as $Q_s$ of the MPM, EH, and Bagnold sediment transport formulas with respect to $Q$, $\Psi$ or $S$ for the three different channel types and  $\alpha$ values.
Remarkably, the Bagnold formula has a negative exponent for $f$ in all but the pipe-flow channel type.
The total transport formula EH gives a slightly stronger dependence on all variables due to the larger exponent on $\tau$ of $\frac{5}{2}$ versus  $\frac{3}{2}$ for the MPM (Equation~\ref{eq:Qs_eq}).
Albeit, the sediment transport response in pipe flow for the Bagnold case is close to EH.
Conversely, the sediment transport proxies $w_g\,\tau_g$ and $w_f\,\tau_f$ used previously scale only as $Q^2$ for pipe-flow, whereas the sediment transport capacity scales at least with $Q^3$.
The exponent for the width scaling $\alpha$ only impacts the relationship between sediment transport and water discharge in the EH relation in any meaningful way.
However, for the value of $\alpha$ around $\frac{1}{3}$ (a value appropriate for most streams), the exponent on $Q$ for EH is only slightly greater than the other transport relations.
